---
title: "Homework 1"
subtitle: "DSCI 101"
author: "Francis Olakangil"
date: "Due 09-11-2023"
output: html_document
---
Homework should be written as an R Markdown file and the knitted output
should be submitted. Homework should be turned in on Sakai.

1. (4pts) Set up a github account and create a repository called DSCI101.
Clone this repo into RStudio: https://github.com/menawhalen/DSCI_101_F23

2. (3pts) Create a numeric vector containing the values 5, 10, 15, 20, and 25. Calculate the mean of this vector and assign it to a variable called `mean_value`. Print the `mean_value`

```{r ex2}
vec <- c(5, 10, 15, 20, 25)
mean_value <- mean(vec)
print(mean_value)
```

3. (3pts) Create a data frame with the following information about three of your friends:

* Name (character)
* Age (numeric)
* Favorite Color (character)

Make sure to include at least three rows of data. Print the entire data frame. Name it whatever you like.

```{r ex3}
friends <- data.frame(names = c("Mike", "Dylan", "Ryan"),
              age = c("16", "17", "14"),
              fav_color = c("blue", "red", "green"))
print(friends)
```

4. (3pts) Create two numeric vectors: `vector1` with values 1, 2, 3, and `vector2` with values 4, 5,
```{r}
vec1 <- c(1, 2, 3)
vec2 <- c(4, 5)
```

6. Perform the following operations:

* Add `vector1` and `vector2` and store the result in a new vector called `sum_vector.`
* Subtract `vector2` from `vector1` and store the result in a new vector called `diff_vector.`
* Multiply `vector1` and `vector2` element-wise and store the result in a new vector called `prod_vector.`
* Divide `vector1` by `vector2` element-wise and store the result in a new vector called `quot_vector.`

Print all four result vectors.

```{r ex4}
sum_vector <- vec1 + vec2
sum_vector

diff_vector <- vec1 - vec2
diff_vector

prod_vector <- vec1 * vec2
prod_vector

quot_vector <- vec1/vec2
quot_vector

print(sum_vector)
print(diff_vector)
print(prod_vector)
print(quot_vector)
```

5. (12pts) Using the `top_colleges_2022` dataset answer the following questions.

a) Read in the csv file and name the variable `college`. This should be simple if question 1 has been completed as the github will download the data folder.

```{r ex5a}
library(tidyverse)
college <- read_csv("../data/top_colleges_2022.csv")
```

b) How many colleges in this data set are in Illinois.

```{r ex5b}
#view(college)
illinois_colleges <- college %>% 
  filter(state...14 == "Illinois")
nrow(illinois_colleges)

```

c) Find the undergraduate student population of LUC. How many colleges (in this dataset) have a larger undergraduate student population than LUC?

```{r ex5c}
#view(college)
luc_pop <- college %>% 
  filter(organizationName == "Loyola University Chicago") %>% 
  pull(undergradPop)
luc_pop

colleges_pop <- college %>% 
  filter(undergradPop > luc_pop)
nrow(colleges_pop)

```