---
title: "Homework 2"
author: "Francis Olakangil"
date: "2023-09-19"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
options(repos = c(CRAN = "https://cran.r-project.org"))
#explicitly setting CRAN mirror for lahman installation to fix lahman pakcage installation error, credit to stackoverflow

```
1. Using the `cubs_all_time` dataset found in the data folder. Answer the following questions using R:

a) (10 points) How many total games have the Cubs won and lost between 1876 and 2022?

```{r}
library(tidyverse)

cubs <- read_csv("~/Library/CloudStorage/OneDrive-LoyolaUniversityChicago/DSCI101/DSCI_Whalen/data/cubs_all_time.csv")
#kept giving me an error when reading in file with relative file path, had to paste in full path
#view(cubs)

win_sum <- cubs %>% 
  summarise(winSum = sum(W))
win_sum

loss_sum <- cubs %>% 
  summarise(lossSum = sum(L))
loss_sum
```

b) (10 points) What year did the Cubs score the most runs? What year did the Cubs score the fewest runs? Do you have any thoughts about the year that the Cubs scored the fewest runs?

```{r ex1b}
most_runs <- cubs %>%
  filter(R == max(R)) %>% 
  select(Year)
most_runs

fewest_runs <- cubs %>% 
  filter(R == min(R)) %>% 
  select(Year)
fewest_runs

#Thoughts about year of fewest runs: This was also the year of the start of the pandemic, which is why the total number of games the Cubs played was significantly reduced. This is probably why such a significantly smaller number of runs was scored across the season.

```

c) (10 points) In how many seasons was the Cubs total attendance (i.e. the variable Attendance) over 3 million?

```{r ex1c}
total_attendance <- cubs %>% 
  filter(Attendance > 3000000)
total_attendance
nrow(total_attendance)


```

d) (10 points) In the years between 1982 and 2000 (inclusive), what was the greatest ratio of runs for (R) to runs against (RA) in a single season?

```{r ex1d}
years_bw <- cubs %>% 
  filter(Year >= 1982 & Year <= 2000)
years_bw

ratio_column <- years_bw %>% 
  mutate(ratioColumn = R / RA)
ratio_column

greatest_ratio <- ratio_column %>% 
  filter(ratioColumn == max(ratioColumn)) %>% 
  pull(ratioColumn)
greatest_ratio

cat("In between these years (inclusive), the greatst ratio of runs for to runs against in one season was", greatest_ratio)

```

2. Using the Teams data frame in the Lahman package:

```{r}
install.packages("Lahman")
#The data is in this package
library(Lahman)
#The data set is called "Teams".
#Load this package and the data will automatically be loaded.
## data("Teams") will put it in your environment
#Use help(Teams) to get a description of the data

data("Teams")
#view(Teams)
help(Teams)
```

a) (10 points) Create a data frame that is a subset of the Teams data frame that contains only the years from 2000 through 2009 and the variables yearID, W, and L.

```{r ex2a}
year_subset <- Teams %>% 
  filter(yearID >= 2000 & yearID <= 2009) %>% 
  select(yearID, W, L)
glimpse(year_subset)

```


b) (10 points) How many years did the Chicago Cubs (teamID is "CHN") hit at least 200 HRs in a season and what was the median number of wins in those seasons.

```{r ex2b}
cubs_wins <- Teams %>% 
  filter(teamID == "CHN" & HR >= 200)
cubs_wins
cat("The number of years that the Cubs hit at least 200 HRs in a season was", nrow(cubs_wins))

median_wins <- cubs_wins %>% 
  summarise(medianWins = median(W))
median_wins

```

3. Using `flights` dataset from `nycflights13` 

```{r}
library(nycflights13)
#The data set is called "flights".
#Load this package and the data will automatically be loaded.
# data("flights") will put it in your environment
#Use help(flights) to get a description of the data
data("flights")
help(flights)

view(flights)
```

Use the nycflights13 package and the flights data frame to answer the following questions (you can ignore missing values):

a) (20 points) What month had the highest average time of delayed flights? Is there a difference between using departure delay versus arrival delay?

```{r ex3a}
dep_delays <- flights %>% 
  group_by(month) %>% 
  summarise(dep_mean = mean(dep_delay, na.rm = TRUE)) %>% 
  arrange(dep_mean)
dep_delays

arr_delays <- flights %>% 
  group_by(month) %>% 
  summarise(arr_mean = mean(arr_delay, na.rm = TRUE)) %>% 
  arrange(arr_mean)
arr_delays

#Although month 7 ranked as having the highest average time of delayed flights across both departures and arrivals, there is a small difference bw using either as metrics, considering the other entries weren't in the exact same order. The delays were longer than arrivals across the board.

```

b) (20 points) Which airport has the highest proportion of delayed flights leaving? Which airport has the highest proportion of flight arriving late?

```{r ex3b}
delay_leave <- flights %>% 
  mutate(delaysLeave = dep_delay > 0) %>% 
  group_by(origin) %>% 
  summarise(total_delays = sum(delaysLeave, na.rm = TRUE)/n()) %>% 
  #dividing by n(), wanting to find proportion by summing delays by leaving and dividing by number of rows.
  arrange(total_delays)
delay_leave
#the airport at the bottom of the list has the highest proportion of delayed flights leaving.

delay_arrival <- flights %>% 
  mutate(arrLeave = arr_delay > 0) %>% 
  group_by(dest) %>% 
  #using dest because measuring the delay that the plane takes to arrive at the specific destination, not the origin of where it started
  summarise(arr_delays = sum(arrLeave, na.rm = TRUE)/n()) %>% 
  arrange(arr_delays)
delay_arrival
#airport at bottom of list has highest proportion of delayed flights arriving at that airport (using arrange() to order that way specifically)

```

