---
title: "Homework 7"
author: "Francis Olakangil"
date: "2023-11-28"
output: html_document
---


```{r}
## Load Gestation data
library(mosaicData)
library(tidyverse)

```


### 1. Find median age of the mother in this data set. Then use bootstrapping to estimate the standard error of the median age of the mother.

```{r}
data(Gestation)
#view(Gestation)
#?Gestation
#standard error: SD divided by sqrt of n

# finding original median, make sure to pipe into Gestation
og_median <- Gestation %>% 
  summarise(median(age, na.rm = TRUE))
print(paste("Median of Mother's age is", og_median))

# bootstrapping function for standard error
# take samples of values in age column, set replacement to TRUE to resample values
bootstrap_fnx <- function(age) {
  resampled_data <- sample(age, replace = TRUE)
  return(median(resampled_data, na.rm = TRUE))
  # fxn also sets up median of new samples for standard error op
}

# resampling for bootstrap
num_samples <- 10000
bootstrap_medians <- map_dbl(1:num_samples, ~bootstrap_fnx(Gestation$age))
# applies btstrp fxn to certain number of samples/attempts

# final standard error calc
bootstrap_SE <- sd(bootstrap_medians)
print(paste("The standard error of the median age of the mother is", bootstrap_SE))
# small SE indicates low variability, probably bc the median, in constrast to the mean, was used
```

### 2. Alcohol assumption is significantly associated with lung cancer. Name a variable(s) that is likely confounding the relationship between alcohol consumption and lung cancer.
One such variable most likely confounding the relationships between alcohol consumption and lung cancer is smoking. When identifying a confounding variable, we are looking for a variable that has some connection to both the independent and dependent variable. In this case, it is likely that those who smoke often also drink. Since smoking has a strong correlation with lung cancer, it confounds the relationship between the cancer and alcohol because the two activities are 1) often done together by the same peoples and 2) smoking causes lung cancer. Another possible variable that confounds the relation between alcohol and lung cancer could be genetics, where someone is genetically predisposed to both contracting lung cancer and becoming an alcoholic. Another possible confounding variable might be one's environment, where one's job might expose them to conditions that exacerbate the chances of contracting lung cancer. Being in these jobs might also infer an income bracket of people who drink more.

Since there may not always be a direct connection between independent and dependent variables in statistical studies, it is important to identify confounding variables that may skew studies and the conclusions that are made from them.

### 3. Explain why overfitting in machine learning is bad.
When we, as data scientists, use machine learning and their associated models to make predictions, we train the model on a given data set. However, doing this can result in the behavior only accurately making predictions for the given data set that it was trained on, while giving inaccurate data for further data sets fed into the model. When a model is overfitted, it is built too closely to the training data set, meaning that it learned all of the small, often meaningless, nuances of the training data set instead of emphasizing the main patterns in it that the data scientists would have wanted it to learn. As a result, the model becomes overfitted, where it will start to use the parameters of the training data set to help it make predictions about the other data sets fed into it. In other words, an overfit machine learning model is unable to generalize to new data, creating poor predictions when applied to other datasets. Ultimately, the main problem with overfitting in machine learning is that you aren't creating a generalized, applicable model that can predict well with data it has never seen before. In order to create a robust model, it needs to theoretically be able to accurately make predictions on data sets it hasn't been trained on - overfitting doesn't preprare the model in that way.



